\documentclass{article}

\usepackage{amssymb,amsmath}

\usepackage{hyperref}
\usepackage[nameinlink]{cleveref}
\usepackage{enumerate}
\usepackage{tikz-cd}
\usepackage{graphicx}

\usepackage{mathrsfs}
%% Fancy fonts --- feel free to remove! %%
\usepackage{Baskervaldx}
\usepackage{mathpazo}


\crefname{section}{Section}{Sections}
\crefname{equation}{}{}

%% Theorem environments %%

\usepackage{amsthm}

  \theoremstyle{plain}

  \newtheorem{innercustomproposition}{Proposition}
  \crefname{innercustomproposition}{Proposition}{Propositions}
  \newenvironment{proposition}[1]
    {\renewcommand\theinnercustomproposition{#1}\innercustomproposition}
    {\endinnercustomproposition}

  \newtheorem{innercustomlemma}{Lemma}
  \crefname{innercustomlemma}{Lemma}{Lemmas}
  \newenvironment{lemma}[1]
    {\renewcommand\theinnercustomlemma{#1}\innercustomlemma}
    {\endinnercustomlemma}

  \newtheorem{innercustomcorollary}{Corollary}
  \crefname{innercustomcorollary}{Corollary}{Corollaries}
  \newenvironment{corollary}[1]
    {\renewcommand\theinnercustomcorollary{#1}\innercustomcorollary}
    {\endinnercustomcorollary}

  \newtheorem*{theorem}{Theorem}


  \theoremstyle{definition}

  \newtheorem*{remark}{Remark}

  \newtheorem{innercustomdefinition}{Definition}
  \crefname{innercustomdefinition}{Definition}{Definitions}
  \newenvironment{definition}[1]
    {\renewcommand\theinnercustomdefinition{#1}\innercustomdefinition}
    {\endinnercustomdefinition}


%% Shortcuts %%

\newcommand{\sh}{\mathscr}
\newcommand{\cat}{\mathcal}

\renewcommand{\geq}{\geqslant}
\renewcommand{\leq}{\leqslant}

\newcommand{\todo}{\textbf{ !TODO! }}
\newcommand{\oldpage}[1]{\marginpar{\footnotesize$\Big\vert$ \textit{p.~#1}}}

\renewcommand{\thepart}{\Alph{part}}

%% Document %%

\usepackage{embedall}
\begin{document}

\renewcommand{\abstractname}{Translator's note.}

\title{Serre's theorem}
\author{P. Gabriel}
\date{}
\maketitle

\begin{abstract}
  \renewcommand*{\thefootnote}{\fnsymbol{footnote}}
  \emph{This text is one of a series\footnote{\url{https://github.com/thosgood/translations}} of translations of various papers into English.}
  \emph{The translator takes full responsibility for any errors introduced in the passage from one language to another, and claims no rights to any of the mathematical content herein.}
  
  \emph{What follows is a translation (last updated \today) of the French paper:}

  \medskip\noindent
  \textsc{Gabriel, P.}
  ``Le th\'{e}or\`{e}me de Serre''.
  \emph{S\'{e}minaire Claude Chevalley}, Volume~\textbf{4} (1958-1959), Talk no.~2, 8~p.
  {\footnotesize\url{http://www.numdam.org/item/SCC_1958-1959__4__A2_0/}}
\end{abstract}

\setcounter{footnote}{0}

\tableofcontents


%% Content %%

\bigskip\bigskip


\part{Affine algebraic sets and classical functors}
\label{chapterA}


\section{Ringed topological spaces}
\label{section1}

\oldpage{2-01}

From now on, and unless otherwise mentioned, the rings we consider will be assumed to be commutative, with a unit element, and \emph{Noetherian}.
We define a \emph{ringed topological space} $(V,\sh{A})$ to be a topological space $V$ endowed with the structure defined by the data of a sheaf of rings $\sh{A}$.
If $(V,\sh{A})$ and $(W,\sh{B})$ are two ringed topological spaces, then we define a morphism from the first to the second by the data of:
\begin{enumerate}[(a)]
  \item a continuous map $\psi\colon V\to W$;
  \item for every open subset $U$ of $W$, a ring homomorphism
    \[
      \varphi_U\colon\sh{B}(U)\to\sh{A}(\varphi^{-1}(U))
    \]
    that is compatible with the restriction maps.
\end{enumerate}

The composition of two morphisms is defined in the evident way, and we speak of the category of ringed topological spaces.
In what follows, $V$ will almost always be a Zariski space.
To such a $V$, we often associate a topological space $S(V)$, which is \emph{the scheme of $V$}, and which is defined in the following way:
\begin{itemize}
  \item the points of $S(V)$ are the closed irreducible subsets of $V$;
  \item the closed subsets of $S(V)$ are the sets $\sh{F}$, where $F$ is a closed subset of $V$, and $\sh{F}$ denotes the set of closed irreducible subsets of $V$ that are contained in $F$.
\end{itemize}

It is then clear that the correspondence $F\mapsto\sh{F}$ between closed subsets of $V$ and closed subsets of $S(V)$ is bijective, that $S(V)$ is a Zariski space, and that every closed irreducible subset of $S(V)$ is the closure of a unique point.
To every sheaf on $V$, we canonically associate a sheaf on $S(V)$.
In particular, if $(V,\sh{A})$ is a ringed topological space, and if $V$ is a Zariski space, then we denote by $(S(V),S(\sh{A}))$ the \emph{scheme of $(V,\sh{A})$}, which is defined to be the ringed topological space given by $S(V)$ and the sheaf associated to $\sh{A}$.
Of course,
\oldpage{2-02}
the scheme of $(S(V),S(\sh{A}))$ is isomorphic to $(S(V),S(\sh{A}))$.

If $A$ is a Noetherian Jacobson ring, and $(\Omega(A),\sh{U})$ is the ringed topological space defined by its maximal spectrum, then the associated scheme is exactly $(V(A),\sh{U})$ (see Talk no.~1).
If $(V,\sh{A})$ is an algebraic set endowed with the sheaf of germs of regular functions, then the associated scheme has been defined by Chevalley in \cite{3} and \cite{4}.
In this case, $V$ is the subspace of $S(V)$ given by the closed points, and $\sh{A}$ is the restriction of $S(\sh{A})$ to $V$.
From now on, we will almost always restrict to the case of algebraic sets over an algebraically closed field.


\section{Sheaves of ideals of an algebraic set}
\label{section2}

If $V$ is an algebraic set, and $\sh{A}$ its sheaf of germs of regular functions, then the notion of quasi-coherent (resp. coherent) sheaves generalises the notion of a module (resp. module of finite type) over the ring of coordinates of an affine algebraic set.
We similarly generalise the notions of support and dimension: if $\sh{F}$ is a coherent algebraic sheaf on $V$, then its \emph{support} is the set of points $x$ of $V$ where the fibre $\sh{F}_x$ of $\sh{F}$ is not zero (if the fibre of $\sh{F}$ is zero at $x$, then it is zero on a neighbourhood of $x$, since $\sh{F}$ is coherent, and thus the support is closed).
We then define the \emph{dimension of $\sh{F}$} to be the dimension of its support.
When $V$ is affine and the coherent sheaf $\sh{F}$ is associated to a module $M$ of finite type, then the support of $\sh{F}$ is given by the set of maximal ideals of the coordinate ring that contain the annihilator of $M$.

The notion of an ideal generalises in the following way: we define a \emph{sheaf of ideals} of $(V,\sh{A})$ to be any quasi-coherent (and thus coherent) subsheaf of $\sh{A}$.
We then find the inevitable correspondence between ideals and closed subsets:

If $\sh{I}$ is a sheaf of ideals, then we associate to it the following closed subset $W(\sh{I})$ of $V$:

$x\in W(\sh{I})$ if and only if $\sh{I}_x$, the fibre of $\sh{I}$ at $x$, is a proper ideal of $\sh{O}_x$ (i.e. $\sh{I}_x\neq\sh{O}_x$), or if and only if the germs of the functions defined by $\sh{I}$ at $x$ are zero.
$W(\sh{I})$ is thus the support of $\sh{A}|\sh{I}$.

Conversely, if $W$ is a closed subset of $V$, then we associate to it the following sheaf of ideals $\sh{I}(W)$:

The sections of $\sh{I}(W)$ on an open subset $U$ are the regular functions defined on $U$ that are zero on $U\cap W$.
If $\sh{I}(U,x)$ denotes the inverse image in $\sh{A}(U)$ of the maximal ideal of $\sh{O}_x$, then $\Gamma(U,\sh{I}(W))$ is the intersection
\oldpage{2-03}
of the $\sh{I}(U,x)$ where $x$ runs over $U\cap W$.
If $U$ is an affine open subset, then the sheaf of ideals $\sh{I}(W)|U$ is exactly the sheaf associated to the ideal $\Gamma(U,\sh{I}(W))$ of $\sh{A}(U)$, which shows that $\sh{I}(W)$ is coherent.

\begin{proposition}{1}
\label{proposition1}
  \begin{enumerate}[(a)]
    \item The map $W\mapsto\sh{I}(W)$ gives a bijective correspondence between the closed subsets of $V$ and the sheaves of ideals $\sh{I}$ that satisfy the following condition:
      for every $x\in V$, $\sh{I}_x$ is either equal to $\sh{O}_x$ or equal to an intersection of prime ideals of $\sh{O}_x$.
    \item The map $W\mapsto\sh{I}(W)$ associates the closed subsets whose connected components are all irreducible to the sheaves of prime ideals (for every $x$, $\sh{I}_x$ is either equal to $\sh{O}_x$ or equal to a prime ideal).
  \end{enumerate}
\end{proposition}

\begin{proof}
  \begin{enumerate}[(a)]
    \item It suffices to give a proof in the case where $(V,\sh{A})$ is an affine algebraic set.
      So let $A=\sh{A}(V)$, and $\mathfrak{a}=\Gamma(V,\sh{I}(W))$.
      We have already seen that $\mathfrak{a}$ is then the intersection of the prime ideals $\sh{I}(V,x)$, where $x$ runs over $W$.
      So $\mathfrak{a}$ is an intersection of prime ideals.
      Since the correspondence between intersections of prime ideals of $A$ and closed subsets of $V$ is bijective, so too is the correspondence between ideals of $A$ and sheaves of ideals of $(V,\sh{A})$;
      it remains only to show, conversely, that the sheaf associated to an intersection $\mathfrak{a}$ of prime ideals satisfies the condition of the proposition: this follows from the conservation properties of the prime decomposition under localisation.
    \item The proof is analogous.
  \end{enumerate}
\end{proof}


\section{The ring of rational functions of an algebraic set}
\label{section3}

We recall that, if $V$ is an algebraic set, then we define a \emph{rational function on $V$} to be a regular map $f$ from an everywhere-dense open subset of $V$ to the field of constants $K$.
We further suppose that the domain of definition of $f$ cannot be extended.
The rational functions on $V$ form an algebra $K(V)$ over $K$.
If we denote by $V_i$ the irreducible components of $V$, then $K(V)$ is isomorphic to the product $\prod_i K(V_i)$, with the isomorphism being the obvious one.
Finally, if $V$ is irreducible, and if $U$ is an affine open subset of $V$, then $K(V)$ is exactly the field of fractions of the coordinate ring of $U$.

The sheaf $\sh{K}(V)$ of rational functions on $V$ is then defined in the following way:
to every open subset $U$ of $V$, we associate the ring $K(U)$ of
\oldpage{2-04}
rational functions on $U$, with the restrictions being obvious.
It is clear that this defines a quasi-coherent sheaf on $V$.
Furthermore, if we denote by $V_i$ the irreducible components of $V$, then $K(U)$ is exactly the product $\prod_{V_i\cap U\neq\varnothing}K(V_i)$.
It thus follows that the sheaf $\sh{K}(V)$ is obtained in the following way:
let $\sh{K}_i$ be the sheaf on $V$ that is zero outside of $V_i$, and that is constant with fibre $K(V_i)$ on $V_i$.
Then $\sh{K}(V)$ is the product of the sheaves $\sh{K}_i$.


\section{Characterisation of affine algebraic sets}
\label{section4}

Let $(V,\sh{A})$ be an algebraic set endowed with its sheaf of rings, and let $A=\Gamma(V,\sh{A})$ be the coordinate ring of $V$.
We know that we then have a canonical morphism $(V,\sh{A})\to(S(V),S(\sh{A}))$, and that $(V,\sh{A})$ is affine if and only if $(S(V),S(\sh{A}))$ is the prime spectrum of an algebra over $K$, of finite type and with no nilpotent elements.

We will now define a morphism $(S(V),S(\sh{A}))\to(V(A),\sh{U})$.
For this, let $x$ be an arbitrary element of $S(V)$, and let $S(\sh{A})_x$ be the fibre of $S(\sh{A})$ at $x$: this an a local ring, and we have a restriction homomorphism $A\to S(\sh{A})_x$.
We denote by $\mathfrak{p}_x$ the prime ideal of $A$ given by the inverse image under this homomorphism of the maximal ideal of $S(\sh{A})_x$.
It is an ideal of functions on $V$ that are zero on the closed point $x$.

The map $\varphi\colon x\mapsto\mathfrak{p}_x$ is a continuous map from $S(V)$ to $V(A)$.
Indeed, it suffices to show that the inverse image of a special open subset $U_f$ of $V(A)$ is an open subset of $S(V)$.
But $\varphi^{-1}(U_f)$ consists of points $x$ of $S(V)$ such that the image $f_x$ of $f$ in $\sh{A}_x$ is invertible, and if $f_x$ is invertible at $x$, then $f_y$ is invertible for $y$ in a neighbourhood of $x$.
QED.

It follows from the above that $1/f$ is a section of $S(\sh{A})$ over $\varphi^{-1}(U_f)$, and so the canonical map from $A$ to $\Gamma(\varphi^{-1}(U_f),S(\sh{A}))$ extends to a homomorphism:
\[
  \varphi_{U_f}\colon A_f \to \Gamma(\varphi^{-1}(U_f), S(\sh{A})).
\]

The $\varphi_{U_f}$ are compatible with the restriction maps, and, by ``passing to the inductive limit'', they define a morphism of ringed topological spaces:
\oldpage{2-05}
\[
  \varphi\colon (S(V),S(\sh{A})) \to (V(A),\sh{U}).
\]

It is clear that $\varphi$ is an isomorphism if and only if $(V,\sh{A})$ is an affine algebraic set.
The following theorem examines this case:

\begin{theorem}[Serre's Theorem]
  The following are equivalent:
\end{theorem}


%% Bibliography %%

\nocite{*}
\bibliographystyle{acm}

\begin{thebibliography}{10}

  \bibitem{1}
  {\sc Borel, A. and Serre, J.-P.}
  \newblock Le th\'{e}or\`{e}m de Riemann-Roch.
  \newblock {\em Bull. Soc. math. France 86\/} (1958), 97--136.

  \bibitem{2}
  {\sc Grothendieck, A.}
  \newblock La th\'{e}orie des classes de Chern.
  \newblock {\em Bull. Soc. math. France 86\/} (1958), 137--154.

  \bibitem{3}
  {\sc S\'{e}minaire Cartan-Chevalley:}
  \newblock G\'{e}om\'{e}trie alg\'{e}brique.
  \newblock {\em Volume~8} (1955/56).

  \bibitem{4}
  {\sc S\'{e}minaire Chevalley:}
  \newblock Classification des groupes de Lie alg\'{e}brique.
  \newblock {\em Volume~1} (1958).

\end{thebibliography}

\end{document}
