\documentclass{article}

\usepackage{amssymb,amsmath}

\usepackage{hyperref}
\usepackage[nameinlink]{cleveref}
\usepackage{enumerate}
\usepackage{tikz-cd}
\usepackage{graphicx}

\usepackage{mathrsfs}
%% Fancy fonts --- feel free to remove! %%
\usepackage{Baskervaldx}
\usepackage{mathpazo}


\crefname{section}{Section}{Sections}
\crefname{equation}{}{}
\crefformat{footnote}{#2\footnotemark[#1]#3}


%% Shortcuts %%

\newcommand{\JJ}{\mathcal{J}}

\renewcommand{\gg}{\mathfrak{g}}
\renewcommand{\geq}{\geqslant}
\renewcommand{\leq}{\leqslant}

\newcommand{\todo}{\textbf{ !TODO! }}
\newcommand{\oldpage}[1]{\marginpar{\footnotesize$\Big\vert$ \textit{p.~#1}}}


%% Document %%

\usepackage{embedall}
\begin{document}

\renewcommand{\abstractname}{Translator's note.}

\title{Gauge transformations in general relativity}
\author{J. Winogradzki}
\date{18\textsuperscript{th} of January, 1955}
\maketitle

\begin{abstract}
  \renewcommand*{\thefootnote}{\fnsymbol{footnote}}
  \emph{This text is one of a series\footnote{\url{https://github.com/thosgood/translations}} of translations of various papers into English.}
  \emph{What follows is a translation (last updated \today) of the French paper:}

  \medskip\noindent
  \textsc{Winogradzki, J}. Les transformations de jauge en relativité généralisée. \emph{Séminaire L. de Broglie. Théories physiques}, Volume~\textbf{24} (1954-1955), Talk no.~10, 10~p. {\footnotesize\url{http://www.numdam.org/item/SLDB_1954-1955__24__A9_0/}}
\end{abstract}

\setcounter{footnote}{0}

\tableofcontents


%% Content %%

\bigskip\bigskip
\oldpage{10-01}
The Universe of the theory of General Relativity is a quadri-dimensional affine and metric space, with the metric being given by a second-order tensor which represents the field.
In general, the affine connection and the metric tensor are asymmetric.
The equations of the field were established in different ways, one of which, due to Einstein, is particularly interesting.
The principle of General Relativity being insufficient to determine the equations of the generalised field, Einstein substituted the principle of General Relativity with a stricter theory of Relativity, and it is from this principle that he deduced the field equations.
He showed that, effectively, the field equations are practically determined if we ask for them to be invariant under $\lambda$-transformations.
\footnote{\label{footnote1}\textsc{A. Einstein}. Generalization of Gravitation Theory. Appendix II of \emph{the Meaning of Relativity}. Princeton, 1953. \emph{(We will use the notation from this paper.)}\\\textsc{A. Einstein}. Extension du group relativiste. Dans \emph{Louis de Broglie Physicien et Penseur}. Albin Michel, 1953.}

It is not necessary to impose a priori the precise nature of the extension of the relativistic group; some simple consideration suggests a postulate from which it can be determined.
The same postulate leads to the field equations.
\footnote{\textsc{J. Winogradzki}, \emph{C.R. Acad. Sc.}, \textbf{239}, 1954, p.~1359.}


\section{Axiomatic foundations}
\label{section1}

One of the classical methods for establishing the pure gravitational field equations consists of deducing them from the set (E) of the following postulates:
\begin{enumerate}[I.]
  \item The universe is a quadri-dimensional affine and metric space, with the metric being given by a second-order tensor.
  \item The affine connection and the metric tensor are symmetric.
  \item The field equations derive from a variational principle.
\oldpage{10-02}
    We vary the affine connection and the metric tensor.
  \item The Hamiltonian is given by the curvature density.
  \item The variations take place without any a priori conditions.
\end{enumerate}

If we remove postulate~II, then the Universe has multiple curvature densities that can be deduced from one another by transposition, either of the metric tensor, or of the affine connection, or both:
\[
  \gg^{ik}R_{ik}, \,\,
  \gg^{ik}R_{ki}, \,\,
  \gg^{ik}\widetilde{R}_{ik}, \,\,
  \gg^{ik}\widetilde{R}_{ki}
\]
with $\gg^ik$ the density of the metric tensor, and $R_{ik}$ and $\widetilde{R}_{ik}$ the two Ricci tensors:
\[
\label{equation1}
  R_{ik} = \Gamma_{ik,m}^m - \Gamma_{im,k}^m + \Gamma_{pm}^m\Gamma_{ik}^p - \Gamma_{ip}^m\Gamma_{mk}^p
  \tag{1}
\]
\[
\label{equation2}
  \widetilde{R}_{ik} = \Gamma_{ki,m}^m - \Gamma_{mi,k}^m + \Gamma_{mp}^m\Gamma_{ki}^p - \Gamma_{pi}^m\Gamma_{km}^p.
  \tag{2}
\]
Avoiding any arbitrary choice, we modify postulate~IV:
\begin{enumerate}
  \item[IVa.] The Hamiltonian is given by any one of the curvature densities.
\end{enumerate}

The set (E'), consisting of postulates~I, III, IVa, and V, does not determine the field equations in an unequivocal way, since the field equations then depend on the choice of the Hamiltonian.
The set (E') of postulates is thus unacceptable.
We will try to modify it whilst staying as close as possible to (E).

The set (E) of postulates is equivalent to the set containing postulates~I, III, IVa, and another postulate that combines II and V:
\begin{enumerate}
  \item[Va.] The variations take places under the a priori conditions $g_{ik}=g_{ki}$ and $\Gamma_{km}^i=\Gamma_{mk}^i$.
\end{enumerate}

So (E') differs from (E) only by the last postulate; it is thus this postulate that we will modify.
We will base our modification on the Einsteinian idea of an extension of the relativistic group\cref{footnote1}, but without imposing a priori, as Einstein did, the precise nature of this extension.

The numerous systems of field equations following from (E') are already invariant, not only under coordinate transformations, but also under certain other transformations.
The hypothesis that the field equations satisfy a stricter principle of Relativity than that of General Relativity is thus not sufficient to indicate how we must modify postulate~V.
So, short of introducing an arbitrary limitation, the only option left is to accept the validity of a principle of Relativity
\oldpage{10-03}
that is as strict as possible, i.e. the invariance of the field equations with respect to the largest group possible.
To give the hypothesis the necessary precision, it suffices to accept that the transformations that constitute the extension of the relativistic group (that we will call ``gauge transformations'', to evoke the analogy with electromagnetism) posses certain general properties.
Denote by $\JJ$ the group of transformations possessing these general properties of gauge transformation.
To obtain the field equations that satisfy the stricter principle of Relativity, all whilst modifying postulate~V as little as possible, we will replace postulate~V by the following:
\begin{enumerate}
  \item[Vb.] The variations take place without any a priori conditions, unless a variation linked to the metric tensor or the affine connection does not give field equations that satisfy a stricter principle of Relativity than the equations obtained by free variation\todo.
    In this case, we impose \todo
\end{enumerate}

If we adopt the natural strong definition of the group $\JJ$ that follows, then the set (E''), consisting of postulates~I, III, IVa, and Vb, determines the extension of the relativistic group (\cref{section2}) and leads to the field equations (\cref{section2,section3}).

It remains to determine the properties of the $\JJ$-transformations, i.e. the general properties to impose on the gauge transformations.
We first of all accept that these transformations, like the gauge transformations of electromagnetism, leave the field invariant, and are independent from it.
Since, in the classical theory of the gravitational field, the metric tensor constitutes the field, it should, in generalised gravity, if not constitute the field, then at least be a part of it.
The $\JJ$-transformations can thus be written as
\[
\label{equation3}
  \begin{cases}
    J(g_{ik}) = g_ik
  \\J(\Gamma_{km}^i) = \Gamma_{km}^i+\Lambda_{km}^i
  \end{cases}
  \tag{3}
\]
where $\Lambda_{km}^i$ is independent of the metric tensor, and is a tensor by virtue of being the difference of two affine connections.
We accept that this tensor does not depend \todo

\end{document}
