\documentclass{article}

\usepackage[margin=1.6in]{geometry}

\title{Summary of essential results in the theory of topological tensor products and nuclear spaces}
\author{A.~Grothendieck}
\date{}

\newcommand{\doctype}{French paper}
\newcommand{\origcit}{%
  \textsc{Grothendieck, Alexander}.
  ``R\'{e}sum\'{e} des r\'{e}sultats essentiels dans la th\'{e}orie des produits tensoriels topologiques et des espaces nucl\'{e}aires''.
  \emph{Annales de l'institut Fourier}, Volume~\textbf{4} (1952) , 73--112.
  {\url{http://www.numdam.org/item/?id=AIF_1952__4__73_0}}%
}


\usepackage{amssymb,amsmath}

\usepackage{hyperref}
\usepackage{xcolor}
\hypersetup{colorlinks,linkcolor={blue!50!black},citecolor={blue!50!black},urlcolor={blue!80!black}}
\usepackage{enumerate}
\usepackage{tikz-cd}

\usepackage{mathrsfs}
%% Fancy fonts --- feel free to remove! %%
\usepackage{fouriernc}


\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{xstring}
\makeatletter
\ifx\pdfmdfivesum\undefined
  \let\pdfmdfivesum\mdfivesum
\fi
\edef\filesum{\pdfmdfivesum file {\jobname}}
\pagestyle{fancy}
\makeatletter
\let\runauthor\@author
\let\runtitle\@title
\makeatother
\fancyhf{}
\lhead{\small\runtitle}
\lfoot{\footnotesize Version: \texttt{\StrMid{\filesum}{1}{8}}}
\cfoot{\small\thepage\ of \pageref*{LastPage}}


%% Theorem environments %%

\usepackage{amsthm}

\theoremstyle{plain}

\newenvironment{itenv}[1]
  {\phantomsection\par\medskip\noindent\textbf{#1.}\itshape}
  {\medskip}

\newenvironment{rmenv}[1]
  {\phantomsection\par\medskip\noindent\textbf{#1.}\rmfamily}
  {\medskip}


%% Shortcuts %%

\newcommand{\aster}[1]{$\star${#1}$\star$}
\newcommand{\BB}{\mathrm{B}}
\newcommand{\LL}{\mathrm{L}}
\newcommand{\JJ}{\mathrm{J}}
\newcommand{\CC}{\mathrm{C}}
\renewcommand{\ll}{\ell}
\newcommand{\sBB}{\mathscr{B}}
\newcommand{\sLL}{\mathscr{L}}
\newcommand{\sEE}{\mathscr{E}}
\newcommand{\DF}{\mbox{\normalfont($\mathscr{DF}$)}}
\newcommand{\FF}{\mbox{\normalfont($\mathscr{F}$)}}
\newcommand{\hotimes}{\widehat{\otimes}}
\newcommand{\hhotimes}{\skew{0}{\widehat}{\skew{0}\widehat{\otimes}}}
\newcommand{\scr}{\mathscr}
\newcommand{\dd}{\mathrm{d}}

\newcommand{\todo}{\textbf{ !TODO! }}
\newcommand{\oldpage}[1]{\marginpar{\footnotesize$\Big\vert$ \textit{p.~#1}}}


%% Document %%

\usepackage{embedall}
\begin{document}

\maketitle
\thispagestyle{fancy}

\renewcommand{\abstractname}{Translator's note.}

\begin{abstract}
  \renewcommand*{\thefootnote}{\fnsymbol{footnote}}
  \emph{This text is one of a series\footnote{\url{https://thosgood.com/translations}} of translations of various papers into English.}
  \emph{The translator takes full responsibility for any errors introduced in the passage from one language to another, and claims no rights to any of the mathematical content herein.}

  \medskip
  
  \emph{What follows is a translation of the \doctype:}

  \medskip\noindent
  \origcit
\end{abstract}

\setcounter{footnote}{0}

\tableofcontents
\bigskip


%% Content %%

\section*{Introduction}
\addcontentsline{toc}{section}{\protect\numberline{}Introduction}%


\subsection*{Subject}
\addcontentsline{toc}{subsection}{\protect\numberline{}Subject}%

\oldpage{73}

\footnote{The numbers in brackets refer to the bibliography found at the end of this article.}
This article aims to give a summary, without proofs, of the principal results found in my work ``Produits tensoriels topologiques et espaces nucl\'{e}aires'', which will be published in the \emph{Memoirs of the Amer. Math. Society} (and which I will refer to as \cite{PTT}).
The main concern throughout \cite{PTT} was that of being exhaustive, both in terms of studying all the questions raised by the topics covered, as well as trying to state the more difficult results as theorems that were as general as possible.
This work was also very dense, and the important simple ideas risked being sometimes hidden by technical details.
This is why this bowdlerised summary is possibly useful in giving a more assimilable outline of the theory.
Some extra comments, interesting but not necessary for the general understanding of this summary, as well as some hints for certain proofs, have been placed between stars, \aster{like this}.

The importance of topological tensor products shows itself in many different settings:
\begin{enumerate}[a)]
  \item The notion of the topological tensor product forms the foundations of a simple and general formulation of \emph{Fredholm theory}, including, alongside the classical case of an integral operator defined by a continuous kernel, many other operators that are defined in the most important functional spaces.%
    \footnote{Such a formulation of Fredholm theory seems to have appeared for the first time in \textsc{A.~Ruston}, ``Direct product of Banach spaces and linear functional equations'', \emph{Proc. of the London Math. Soc.} \textbf{3} (1951), 1. My work on this subject was conceived independently of his (in the autumn of 1951), and is rather different.}
  \item The many variants of the notion of topological tensor product give rise, by duality, to the definition of many remarkable classes of bilinear forms and linear operators, whose
\oldpage{74}
    study is only just barely covered in \cite[chap.~I, \S4]{PTT}.
    In particular, the techniques introduced there, conveniently systematised and exploited, allow us to obtain entirely unexpected results in the \emph{theory of linear transformations between the space $L^1$, $L^2$, and $L^\infty$}, and their topological-vectorial analogues (these results being, as of yet, not definitive, and thus unpublished).
    I might return to this subject, and restrict myself to explaining, in a rather different way, the systematic work of von Neumann--Schatten on the remarkable classes of compact operators in a Hilbert space \cite[chap.~4]{8}.
  \item From the point of view of this current work, the most important application of topological tensor products is the theory of \emph{nuclear spaces}.
    We explain this theory, generalise it, and make precise the famous ``theory of kernels'' of L.~Schwartz, and further discover new properties, even for the most classical of spaces.
    Here the topological tensor calculus is the most simple, since the majority of variants of the notion of topological tensor product coincide, and their properties thus sum.
    For now, there are not many applications to particular theories of the general theorems that we obtain.
    The most interesting seems to be a topological-vectorial variant of the ``K\"{u}nneth theorem'', giving the homology of a complex defined as the tensor product of two complexes, a variant which seems useful in topological algebra.
  \item Generally, it seems to me that the notions of topological tensor product are perfect for giving a suggestive and manageable \emph{language}, that would be good to use in many situations in functional analysis, especially since we have theorems (some of which are non-trivial) at our disposition that we can benefit from.
  I hope that this summary, or, better, \cite{PTT}, will succeed in giving the reader a similar impression, before the publication of the articles promised above.
\end{enumerate}


\subsection*{Terminology and notation}
\addcontentsline{toc}{subsection}{\protect\numberline{}Terminology and notation}%

Generally we follow the terminology and notation of \cite{3}, apart from the fact that we call the semi-reflexive spaces of \cite{3} \emph{reflexive}.
We only consider, unless otherwise stated, spaces that are \emph{locally convex} and \emph{separated};
by ``quotient space'' of a space $E$, we mean the quotient of $E$ by a \emph{closed} vector subspace.
The \emph{dual} of $E$, written $E'$, is assumed to be, unless otherwise stated, endowed with the strong topology (i.e. the topology
\oldpage{75}
of bounded convergence).
The dual of $E'$, or the \emph{bidual} of $E$, written $E''$, is assumed to be, unless otherwise stated, endowed with the topology given by uniform convergence on the equicontinuous subsets of $E'$, which induces the original topology on $E$.
We will eventually need to appeal to certain notions defined and studied in \cite{6}, most notably that of a \emph{$\DF$-space}.
For our purposes here, it will suffice to know that the dual of a $\FF$-space is a $\DF$-space; that every normed space is a $\DF$-space; and that the dual of a $\DF$-space is an $\FF$-space.

Let $E$, $F$, and $G$ be locally convex spaces.
Denote by $\BB(E,F;G)$ (resp. $\sBB(E,F;G)$) the space of continuous bilinear maps (resp. of separately continuous bilinear maps, i.e. linear with respect to each variable) from $E\times F$ to $G$.
Denote by $\LL(E;F)$ the space of continuous linear maps from $E$ to $F$.
Denote by $\sBB_e(E'_s,F'_s)$ the space of separately continuous bilinear forms on the product of the weak duals $E'_s$ and $F'_s$ of $E$ and $F$, endowed with the \emph{biequicontinuous topology} topology, i.e. the topology given by uniform convergence on the products of an equicontinuous subset of $E'$ with equicontinuous subset of $F'$.
This space is complete if and only if the spaces $E$ and $F$ are complete.

We define a \emph{bounded} (resp. \emph{compact}, resp. \emph{weakly compact}) \emph{linear map} from $E$ to $F$ to be a linear map from $E$ to $F$ that sends a suitable neighbourhood of $0$ to a bounded (resp. relatively compact, resp. relatively weakly compact) subset of $F$.

For short\footnote{This terminology was suggested to me by R.E.~Edwards.\\\emph{[Trans.] The seemingly more popular terminology these days is to say ``absolutely convex'' instead of ``disked'', and to speak of the ``absolutely convex hull'' instead of the ``disked hull''.}}, if $E$ is a vector space, we say \emph{disk} or \emph{disked set in $E$} to mean a convex and circled (a.k.a. balanced) subset of $E$.
If $E$ is a locally convex space, and $A$ a bounded disk in $E$, then we denote by $E_A$ the vector space generated by $A$, and endowed with the norm $\|x\|_A=\inf_{x\in\lambda A}|\lambda|$.
If $A$ is closed, then the unit ball of $E_A$ is $A$.
If $A$ is complete, then $E_A$ is complete.
If $V$ is a disked neighbourhood of $0$ in $E$, then $E_V$ denotes the normed space given by passing to the quotient under the semi-norm $\|x\|_V=\inf_{x\in\lambda V}|\lambda|$.

Recall that a locally convex space is said to be \emph{quasi-complete} if its closed bounded subsets are complete, \emph{barrelled} (resp. \emph{quasi-barrelled}) if the bounded subsets of its weak dual (resp. of its strong dual) are equicontinuous, and \emph{bornological} if every set of linear forms on $E$ that are uniformly bounded on every bounded subset is equicontinuous.
If $E$ is quasi-complete, then tunnelled is equivalent to quasi-tunnelled;
in any case, bornological implies quasi-tunnelled.



\section{Topological tensor products}
\label{1}
\oldpage{76}


\subsection{Generalities on \texorpdfstring{$E\hotimes F$}{E(x)F}}
\label{1.1}

(\cite[chap.~1, \S1, n\textsuperscript{o}~1 and n\textsuperscript{o}~3]{PTT}).
\medskip

The axiomatic definition of the algebraic tensor product $E\otimes F$ of two vector spaces $E$ and $F$, and of the canonical bilinear map $(x,y)\mapsto x\otimes y$ from $E\times F$ to $E\otimes F$ (\cite{1}) asks only that, for every vector space $G$, the bilinear maps from $E\times F$ to $G$ correspond bijectively with linear maps $f$ from $E\otimes F$ to $G$, where the map corresponding to $f$ is given by $(x,y)\mapsto f(x\otimes y)$.

\begin{itenv}{Theorem 1}
\label{1.theorem1}
  If $E$ and $F$ are locally convex spaces, then we can endow $E\otimes F$ with a locally convex topology such that, for every locally convex space $G$, the \emph{continuous} bilinear maps from $E\times F$ to $G$ correspond exactly to \emph{continuous} linear maps from $E\otimes F$ to $G$.
  Further, such a topology is unique.
\end{itenv}

Then the \emph{equicontinuous} subsets of $\BB(E,F;G)$ correspond exactly to the \emph{equicontinuous} subsets of $\LL(E\otimes F;G)$ as well.
Unless otherwise mentioned, $E\otimes F$ is assumed to be endowed with the above topology, called the \emph{projective tensor product of the topologies of $E$ and $F$};
endowed with this topology, $E\otimes F$ is called the \emph{projective topological tensor product} of $E$ and $F$.

If $E$ and $F$ are normed, then $E\otimes F$ is normable, and we can even find a norm such that, for every \emph{normed} space $G$, the above isomorphism between $\BB(E,F;G)$ and $\LL(E\otimes F;G)$ preserves the natural norms.
Further, such a norm is unique.
This norm on $E\otimes F$, denoted by $u\mapsto\|u\|$, where the norms of $E$ and $F$ are implicit, is the lower bound of the quantities $\sum_i\|x_i\|\|y_i\|$ over all representations of $u$ in the form $u=\sum_i x_i\otimes y_i$ (and this norm has already been considered, in \cite{8}).
It is also the gauge of the set $\Gamma(U\otimes V)$, where $U$ (resp. $V$) is the unit ball in $E$ (resp. $F$), and where $U\otimes V$ denotes the set of $x\otimes y$ such that $x\in U$ and $y\in V$
\oldpage{77}
(with $\Gamma$ denoting, as per usual, the balanced hull).
In the case where $E$ and $F$ are general locally convex spaces, a fundamental system of neighbourhoods of $0$ in $E\otimes F$ is obtained by taking the sets $\Gamma(U\otimes V)$, where $U$ (resp. $V$) runs over a fundamental system of neighbourhoods of $0$ in $E$ (resp. $F$).

We can introduce the completion of $E\otimes F$, denoted by $E\hotimes F$, and called the \emph{completed projective tensor product} of $E$ and $F$.
If $E$ and $F$ are normed spaces, then $E\hotimes F$ is a Banach space (with a well-defined norm!).
If $E$ and $F$ are metrisable, then $E\hotimes F$ is of type~$\FF$.
We have, by definition, the \emph{scholium}: if $E$ and $F$ are locally convex spaces, and $G$ is a \emph{complete} locally convex space, then the continuous bilinear maps from $E\times F$ to $G$ correspond bijectively to the continuous linear maps from $E\hotimes F$ to $G$.

This claim still holds true for equicontinuous sets of maps.
In particular, the dual of $E\hotimes F$ is $\BB(E,F)$, with a correspondence between the equicontinuous subsets (which already suffices to characterise the induced topology on $E\otimes F$).

I do not know if, when $E$ and $F$ are of type~$\FF$, this algebraic isomorphism from the dual of $E\hotimes F$ to $\BB(E,F)$ is a \emph{topological} isomorphism, when we endow $\BB(E,F)$ with the topology given by bibounded convergence, i.e. uniform convergence on the products of two bounded sets (``the problem of topologies'').
An equivalent question is the following:
is every bounded subset of $E\hotimes F$ contained in the closed disked hull of a set $A\hotimes B$, where $A$ (resp. $B$) is a bounded subset of $E$ (resp. $F$)?

\aster{
  We now give some general tips for calculations involving $E\hotimes F$ (\cite[chap.~I, \S1, n\textsuperscript{o}~3]{PTT}).
  If $E=\prod_i E_i$ and $F=\prod_j F_j$ (as topological-vectorial products), then $E\hotimes F$ can be identified with $\prod_{i,j}E_i\hotimes F_j$.
  If $E=\sum_i E_i$ (as a topological direct sum), and if $F$ is a normable space, then $E\hotimes F$ can be identified with the topological direct sum $\sum_i(E_i\hotimes F)$.
  This remains true if $F$ is an arbitrary $\DF$~space, provided that $I$ is countable, and these statements can also be generalised to the case where $E$ is the inductive limit (in the most general sense) of a family $E_i$ of spaces.
  If $E$ and $F$ are both of type~$\FF$ (resp. type~$\DF$), then so too is $E\hotimes F$.
  Similarly, if $E$ and $F$ are quasi-normable spaces, or Schwartz spaces (see definitions in \cite[\S3]{6}), then so too is $E\hotimes F$.
}

\oldpage{78}
\subsection{The space \texorpdfstring{$E\hotimes F$}{E(x)F} when \texorpdfstring{$E$}{E} and \texorpdfstring{$F$}{F} are of type \texorpdfstring{$\FF$}{(F)}}
\label{1.2}

(\cite[chap.~1, \S2, n\textsuperscript{o}~1]{PTT}).
\medskip

\begin{itenv}{Theorem 2}
\label{1.theorem2}
  Let $E$ and $F$ be two $\FF$~spaces.
  Then every element of $E\hotimes F$ is the sum of an absolutely convergent series of the form
  \[
    u = \sum_i \lambda_i x_i \otimes y_i
  \]
  where $(x_i)$ (resp. $(y_i)$) is a bounded sequence in $E$ (resp. $F$), and where $(\lambda_i)$ is a summable sequence of scalars.
\end{itenv}

(It is also true that, if $(x_i)$, $(y_i)$, and $(\lambda_i)$ are given as above, then the series $\sum_i\lambda_i x_i\otimes y_i$ is always absolutely convergent in $E\hotimes F$, and so we have a \emph{characterisation} of the elements of $E\hotimes F$).
If $E$ and $F$ are normed, then we can suppose in the above that $\|x_i\|\leq1$, $\|y_i\|\leq1$, $\sum_i|\lambda_i|\leq\|u\|_1+\varepsilon$, where $\varepsilon>0$ is arbitrary and given in advance.
In these two statements, if $u$ runs over a compact subset of $E\hotimes F$, then we can suppose that the sequences $(x_i)$ and $(y_i)$ remain fixed (and we can even suppose that they are sequences that tend to $0$), and that $(\lambda_i)$ runs over a compact subset of $\ll^1$ (the space of summable sequences).
We have an analogous statement for the concrete representation of convergent sequences in $E\hotimes F$.
\hyperref[1.theorem2]{Theorem~2} and its previous variants serve mainly as a way to:
\begin{itenv}{Corollary}
  Let $E$ and $F$ be spaces of type~$\FF$.
  Then every compact subset $K$ of $E\hotimes F$ is contained inside the canonical image of the unit ball of a space $E_A\hotimes F_B$, where $A$ (resp. $B$) is a compact disked subset of $E$ (resp. $F$).
  A fortiori, $K$ is contained in the closed convex hull of $A\otimes B$.
\end{itenv}

This fact also implies that, on $\BB(E,F)$, the ``bicompact convergence'' topology is identical to the compact convergence topology on the dual of $E\hotimes F$.

\aster{
  For the proof of \hyperref[1.theorem2]{Theorem~2}, we suppose, for simplicity, that $E$ and $F$ are Banach spaces, and let $I$ be the product of their unit balls, and $i\mapsto x_i$ and $i\mapsto y_i$ the projections from $I$ to its factors.
  It is easy t see that the linear map $(\lambda_i)\mapsto\sum_i\lambda_i x_i\otimes y_i$ from $\ll^1(I)$ to $E\hotimes F$ is a metric homomorphism from the former to a dense subspace of the latter, and thus \emph{onto} the latter, whence it indeed follows that $E\hotimes F$ can be identified with a quotient space of $\ll^1(I)$.

  From \hyperref[1.theorem2]{Theorem~2}, we can extract results of the following type:
  let $\scr{G}$ be a
\oldpage{79}
  locally compact group (resp. a Lie group), then every summable function $f$ (resp. every infinitely differentiable function of compact support) on $\scr{G}$ is of the form $\sum\lambda_i g_i*h_i$, where $(\lambda_i)\in\ll^1$, and where $(g_i)$ and $(h_i)$ are bounded sequences in $\LL^i(\scr{G})$ (resp. in $\scr{D}(\scr{G})$, the space of infinitely differentiable functions of compact support on $\scr{G}$);
  we thus immediately conclude that $f$ is a linear combination of functions of positive type that are in $\LL(\scr{G})$ (resp. $\scr{D}(\scr{G})$).
  In the former case, we can also restrict to functions that are all of compact support (with the supports of $g_i$ and $h_i$ being contained inside a compact subset that depends only on the compact support of $f$).
  There is a simple direct proof in the case of $\LL^1(\scr{G})$, but I do not think that there is one for $\scr{D}(\scr{G})$, where the question presents difficulties even for $\scr{G}=\mathbb{R}$ (by using the Fourier transformation).
}


\subsection{Calculation of \texorpdfstring{$\LL^1\hotimes E$}{L1(x)E}}
\label{1.3}

(\cite[chap.~I, \S2, n\textsuperscript{o}~2]{PTT}).
\medskip

Let $M$ be a locally compact space endowed with a measure $\mu\geq0$, $E$ a Banach space, and $\LL_E^1(\mu)$ the space of $\mu$-integrable maps from $M$ to $E$ (\cite{2}) endowed with its usual norm: $\|f\|_1=\int\|f(t)\|\dd\mu(t)$.
We denote by $\LL^1(\mu)$ the space of $\mu$-summable scalar functions.
Then there exists an obvious bilinear map $(\varphi,a)\mapsto\varphi\cdot a$ from $\LL^1(\mu)\times E$ to $\LL_E^1(\mu)$ that is of norm~$\leq1$, and thus defines a linear map of norm~$\leq1$ from $\LL^1(\mu)\hotimes E$ to $\LL_E^1(\mu)$.

\begin{itenv}{Theorem 3}
\label{1.theorem3}
  The above map from $\LL^1(\mu)\hotimes E$ to $\LL_E^1(\mu)$ is a metric isomorphic from the former space to the latter.
\end{itenv}

To see this, we can immediately reduce to the case where $E$ is of finite dimension, and then proceed by transposition.
It then suffices to apply the classical theorem of Dunford--Pettis that characterises continuous linear maps from $\LL^1(\mu)$ to $E'$.

If $E$ is an arbitrary locally convex space, then we denote by $\LL_E^1(\mu)$ the completion of the separated space associated to the space of continuous maps with compact support from $M$ to $E$, endowed with its family of semi-norms $f\mapsto\int p(f(t))\dd\mu(t)$ (where $p$ runs over a fundamental family of continuous semi-norms on $E$).
Then \hyperref[1.theorem3]{Theorem~3} easily implies that $\LL_E^1(\mu)$ is again isomorphic to $\LL^1(\mu)\hotimes E$.

\begin{itenv}{Corollary}
  If $E$ is a closed vector subspace of the Banach space $F$, then the canonical linear map from $\LL^1(\mu)\hotimes E$ to $\LL^1(\mu)\hotimes F$ is a metric isomorphism.
\end{itenv}

\oldpage{80}
This recovers, for example, the well-known fact that every continuous linear map from $E$ to the dual $\LL^\infty(\mu)$ of $\LL^1(\mu)$ can be extended to a linear map of the same norm from $F$ to $\LL^\infty(\mu)$;
or, dually, that every continuous linear map from $\LL^1(\mu)$ to a quotient space $F'/E^0$ of a Banach dual by a weakly closed vector subspace comes from a linear map of the same norm from $\LL^1(\mu)$ to $F'$.

\aster{
  The analogue of \hyperref[1.theorem3]{Theorem~3} for $\LL^p$ spaces is false for all $p>1$.
  However, \hyperref[1.theorem3]{Theorem~3} applies, in an essential way, in many important places in the theory outlined here.
  We now give some less important applications (see \cite[chap.~1, \S2, n\textsuperscript{o}~2]{PTT} for details).
  In \hyperref[1.theorem3]{Theorem~3}, if we take $E=c_0$ (the space of scalar sequences that converge to $0$), and note that $\LL_E^1(\mu)$ can then be identified with the space of latticially bounded sequences in $\LL^1(\mu)$ that converge to $0$ almost everywhere, then we see that such sequences in $\LL^1(\mu)$ form a category of invariant sequences from the topological-vectorial point of view.
  In particular, a continuous linear map from $\LL^1(\mu)$ to a space $\LL^1(\nu)$ sends latticially bounded sequences that converge to $0$ almost everywhere to sequences of the same type.
  We thus also see that the latticially bounded subsets of $\LL^1(\mu)$ form an invariant category from the topological-vectorial point of view.
  If we take $F=\ll^p$, with $1\leq p<+\infty$, then we similarly obtain a topological-vectorial interpretation of the sequences $(f_i)$ in $\LL^1(\mu)$ such that
  \[
    \int \left(
      \sum_i |f_i(t)|^p
    \right)^{1/p}
    \dd\mu(t) < +\infty.
  \]
  Such sequences are sent to sequences of the same type by any continuous linear map from $\LL^1(\mu)$ to a space $\LL^1(\nu)$.
  Another interesting application of \hyperref[1.theorem3]{Theorem~3} is the following: every bounded subset $M$ of $\LL^1(\mu)\hotimes E$ is contained in the canonical image of the unit ball of a space $\LL^1(\mu)\hotimes E_A$, where $A$ is a closed bounded disk in the space $E$ of type $\FF$;
  a fortiori, $M$ is contained in the closed disked hull of $B\otimes A$, where $B$ is the unit ball of $\LL^1(\mu)$, which solves the ``problem of topologies'' described in \hyperref[1.1]{\S1.1}.
}


\subsection{Other examples}
\label{1.4}

If $H$ is a Hilbert space, then the elements of $H'\hotimes H$, identified with endomorphisms of $H$ (the \emph{Fredholm maps} or \emph{nuclear maps} from $H$ to $H$ --- see \todo) are exactly the endomorphisms $u$ such that the positive Hermitian operator
\oldpage{81}
$\sqrt{u^*u}$ is compact and has a summable sequence of eigenvalues, and $\|u\|_1$ is then equal to the sum of the eigenvalues of $\sqrt{u^*u}$ (repeated with multiplicities, of course).
We obtain the operators that have already been studied in \cite{4} and \cite{8}.
In fact, $u$ is also a Fredholm operator if and only if its Hermitian components $\frac12(u+u^*)$ and $\frac{1}{2i}(u-u^*)$ are, i.e. if they are compact Hermitian operators whose sequences of eigenvalues are summable.
\emph{Relation to Hilbert--Schmidt operators: If $A$ and $B$ are Hilbert--Schmidt operators, then $AB$ is a Fredholm operator, and $\|AB\|_1\leq\|A\|_2\|B\|_2$};
and, conversely, every Fredholm operator $u$ is the product of two Hilbert--Schmidt operators $A$ and $B$ of norm $\|A\|_2=\|B\|_2=\sqrt{\|u\|_1}$.
All of these facts are elementary (one we know the spectral decomposition of compact Hermitian operators to a Hilbert space) and well known.

Numerous other examples of products $E\hotimes F$, relating to \emph{nuclear spaces}, will be seen in \todo.

\aster{
  In the setting of infinite-dimensional Banach spaces, I do now know, even in particular cases, other concrete characterisations of the elements of $E\hotimes F$ apart from those that we have given.
  Also, the elements of $c_0\hotimes E$ (where $E$ can be any complete locally convex space) can be identified with \emph{certain} sequences in $E$ that converge to $0$, that we can call \emph{nuclearly convergent to $0$};
  but if $E$ is an infinite-dimensional Banach space, then we always obtain a strictly smaller class than the class of all sequences converging to $0$ (see \todo).
  We even show that (if $E$ is an infinite-dimensional Banach space), for any sequence $(\lambda_i)$ of positive scalars that is not square-summable, there exists a sequence $(x_i)$ in $E$ that does not converge nuclearly to $0$, and such that $\|x_i\|=\lambda_i$ for all $i$.
  However, if, for example, $E$ is the space $C(K)$ of continuous functions on a compact space, then we show that every square-summable sequence in $C(K)$ converges nuclearly to $0$.
  We also point out that, in any complete locally convex space, every summable sequence converges nuclearly to $0$.
}


\subsection{\texorpdfstring{$E\hhotimes F$}{E(x)F} spaces}
\label{1.5}

(\cite[chap.~1, \S3, n\textsuperscript{o}~3]{PTT}).
\medskip

If $E$ and $F$ are Banach spaces, then $E\otimes F$ can be considered as a vector subspace of the Banach space $\BB(E',F')$ of continuous bilinear forms
\oldpage{82}
on $E'\times F'$.
The completion of $E\otimes F$ under the norm induced by $\BB(E',F')$ is denoted by $E\hhotimes F$, which is thus a complete normed vector subspace of $\BB(E',F')$.
Every reasonable normed topology on $E\otimes F$ is included between the topology induced by $E\hotimes F$ and the topology induced by $E\hhotimes F$.
If $E$ and $F$ are now both arbitrary locally convex spaces, then we can again consider $E\otimes F$ is a space of bilinear forms on $E'\times F'$, and endow it with the biequicontinuous-convergence topology (i.e. the topology given by uniform convergence on the products of an equicontinuous subset of $E'$ with an equicontinuous subset of $F'$):
the completion of $E\otimes F$ under this topology is again denoted by $E\hhotimes F$.
If $E$ and $F$ are complete, then the space $\sLL_e(E'_s,F'_s)$ of separately continuous bilinear forms on the product $E'_s\times F'_s$ of the weak duals $E'_s$ and $F'_s$ endowed with the biequicontinuous-convergence topology is complete, and so $E\hhotimes F$ can be identified with a topological vector subspace of $\sLL_e(E'_s,F'_s)$.
Then the elements of $E\hhotimes F$ can be identified with certain separately weakly continuous bilinear maps on $E'\times F'$, or even with certain weakly continuous linear maps from $E'$ to $F$;
these linear maps send equicontinuous subsets of $E'$ to relatively compact subsets of $F$, and the converse is true in all known cases (see \todo).

The topology on $E\otimes F$ induced by $E\hotimes F$ is finer than that induced by $E\hhotimes F$, whence we have a canonical continuous linear map
\[
  E\hotimes F \to E\hhotimes F.
\]

An important, unsolved, problem is the question of whether or not this map is always bijective (see \todo).
We note that it seems extremely plausible that, if $E$ and $F$ are Banach spaces such that the above map $E\hotimes F\to E\hhotimes F$ is a topological isomorphism (or even simply a topological homomorphism, i.e. here a map from the first space \emph{onto} the second), then $E$ or $F$ is of finite dimension.
This is true if, for example, $E$ contains a vector subspace isomorphic to an $\ll^p$ space or to $c_0$.

Let $E$ and $F$ be Banach spaces.
The norm induced by the dual of $E\hotimes F$ on $E'\otimes F'$ is clearly the norm induced by $E'\hhotimes F'$.
On the other hand, the norm induced on $E'\otimes F'$ by the dual of $E\hhotimes F$ is, in all known cases (see \todo), identical to the norm induced by $E'\hotimes F'$.
This duality, which barely appears in the present summary, is a precious tool for various questions
\oldpage{83}
(e.g. \cite[chap.~I, \S4, n\textsuperscript{o}~6]{PTT}).
The following theorem (which is, truth be told, trivial), can be seen as the dual response to \hyperref[1.theorem3]{Theorem~3}.

\begin{itenv}{Theorem 4}
\label{1.theorem4}
  Let $M$ be a locally compact space, let $\CC_0(M)$ the space of continuous scalar functions on $M$ that are ``zero at infinity'', endowed with the uniform convergence norm, and let $E$ be a locally convex complete space.
  Then $\CC_0(M)\hhotimes E$ is canonically isomorphic to the space $\CC_0(M,E)$ of continuous maps from $M$ to $E$ that are zero at infinity, endowed with the uniform convergence topology (and with the uniform norm if $E$ is a Banach space).
  In particular, $c_0\hhotimes E$ is isomorphic to the space of sequences in $E$ that tend to $0$.
\end{itenv}

As with the spaces $C(K)\hotimes E$ below, here the spaces $\LL'(\mu)\hhotimes E$ do not, in general, have a special interpretation as functional spaces.
We note, however, that \emph{the space $\ll^1\hhotimes E$ can be understood as the space of summable sequences in $E$} (i.e. of ``commutatively convergent'' sequences in $E$).


\subsection{Tensor product of linear maps}
\label{1.6}

(\cite[chap.~1, \S1, n\textsuperscript{o}~2]{PTT}).
\medskip

Let $E_i$ and $F_i$ ($i=1,2$) be locally convex spaces, and let $u_i$ be a continuous linear map from $E_i$ to $F_i$.
In algebra, we define a linear map $u_1\otimes u_2$ from $E_1\otimes E_2$ to $F_1\otimes F_2$ by the formula $(u_1\otimes u_2)(x_1\otimes x_2)=u_1x_1\otimes u_2x_2$.
This map is continuous if we endow $E_1\otimes E_2$ and $F_1\otimes F_2$ with the topologies induced by $E_1\hotimes E_2$ and $F_1\hotimes F_2$ (resp. by $E_1\hhotimes E_2$ and $F_1\hhotimes F_2$).
It then follows that $u_1\otimes u_2$ extends, by continuity, to a continuous linear map $u_1\hotimes u_2$ from $E_1\hotimes E_2$ to $F_1\hotimes F_2$ (resp. $u_1\hhotimes u_2$ from $E_1\hhotimes E_2$ to $F_1\hhotimes F_2$).
These maps will also simply be denoted by $u_1\otimes u_2$ whenever there is no fear of confusion.

\begin{itenv}{Theorem 5}
\label{1.theorem5}
  If each $u_i$ is a topological homomorphism from $E_i$ to a dense subspace of $F_i$, then $u_1\hotimes u_2$ is a topological homomorphism from $E_1\hotimes E_2$ to a dense subspace of $F_1\hotimes F_2$.
  If each $u_i$ is a topological isomorphism from $E_i$ to $F_i$, then $u_1\hhotimes u_2$ is a topological isomorphism from $E_1\hhotimes E_2$ to $F_1\hhotimes F_2$.
\end{itenv}

These claims remain true if the $E_i$ and $F_i$ are normed and we consider \emph{metric} homomorphisms and \emph{metric} isomorphisms.
Particularly interesting is the following corollary, which can also be obtained by an application of \hyperref[1.theorem2]{Theorem~2}.

\begin{itenv}{Corollary}
  If the $E_i$ and $F_i$ are $\FF$-spaces ($i=1,2$), and the $u_i$ are
\oldpage{84}
  homomorphisms from $E_i$ \emph{onto} $F_i$, then $u_1\hotimes u_2$ is a homomorphism from $E_1\hotimes E_2$ \emph{onto} $F_1\hotimes F_2$.
\end{itenv}

As particular cases of this corollary, we obtain interesting lifting properties of vector functions with values in a quotient space of an $\FF$-space $E$.
If e.g. $f$ is an infinitely differentiable map from an open subset of $\mathbb{R}^n$ to $E/F$, then it comes from an infinitely differentiable map from this same open subset to $E$.
There is an analogous result for holomorphic functions, or infinitely differentiable functions on $\mathbb{R}^n$ that decay rapidly, or summable functions of a certain measure, etc.\footnote{%
  With respect to this point, we done that we can prove, by a completely different method, the following claim, which can be thought of as dual to the corollary of \hyperref[1.theorem3]{Theorem~3}:
  \emph{Let $M$ be a locally compact and paracompact space (e.g. a compact space), and $f$ a continuous map from $M$ to a quotient space $E/F$ of an $\FF$-space $E$. Then $f$ comes from a continuous map from $M$ to $E$.}
  Since the space $\sEE^{(m)}(V)$ of $m$-times continuously differentiable functions on a infinitely differentiable paracompact manifold $V$ is isomorphic to a direct factor of a space of the form $C(M)$ (as I noted in ``Sur les applications lin\'{e}aires faiblement compacts d'espaces du type $C(K)$'', \emph{Can. J. Math.} \textbf{5} (1953), p.~144), it thus follows that the analogous lifting theorem holds also for $m$-times continuously differentiable maps from $V$ to $E/F$.
}
Another application is the following: let $D$ be a differential operator in the space $\sEE(U)$ of infinitely differentiable functions on an open subset $U$ of $\mathbb{R}^n$, let $E$ be an $\FF$-space, and let $\sEE(U,E)$ be the space of infinitely differentiable maps from $U$ to $E$.
Then we have $\sEE(U,E)=\sEE(U)\hotimes E=\sEE(U)\hhotimes E$ (see \hyperref[2.5]{\S2.5}).
Let $D_E$ be the operator in $\sEE(U,E)$ defined by $D$, and then $D_E=D\hotimes1$, where $1$ is the identity on $E$.
Then, if $D$ is a topological homomorphism (resp. an \emph{onto} topological homomorphism), then so too is $D_E$.
Indeed, in the case of an onto homomorphism, this is a particular case of the corollary of \hyperref[1.theorem5]{Theorem~5}, and, in the general case, we use \hyperref[1.theorem5]{Theorem~5} as well as the fact that, for every quotient space $F$ of $\sEE(U)$, we have $F\hotimes E=F\hhotimes E$ (since $\sEE(U)$, and thus $F$, are nuclear; see \hyperref[2.definition1] in \hyperref[2.2]{\S2.2} and \hyperref[2.theorem3]{Theorem~3} in \hyperref[2.3]{\S2.3}).

We note that, if $u_1$ and $u_2$ are topological isomorphisms, then $u_1\hotimes u_2$ is not, in general, a topological isomorphism (nor is, in general $u_1\hhotimes u_2$ a topological homomorphism, if $u_1$ and $u_2$ are \emph{onto} topological homomorphisms).
If each $E_i$ is identified with a topological-vectorial subspace of $F_i$ by $u_i$, then the canonical map $u_1\hotimes u_2$ from $E_1\hotimes E_2$ to $F_1\hotimes F_2$ is a topological isomorphism if and only if every equicontinuous set of bilinear forms on $E_1\times E_2$ is the set of restrictions of an equicontinuous set of bilinear forms on $F_1\times F_2$.
If $F_1$
\oldpage{85}
and $F_2$ are of type~$\FF$, then it suffices to consider sets consisting of \emph{one single} bilinear form.
In general, this criterion will not hold true, but it is linked to an existence problem of topological complements.

\aster{
  More precisely, if $E_1$ and $E_2$ are direct factors, then $E_1\hotimes E_2$ can be identified with a topological-vectorial subspace of $F_1\hotimes F_2$.
  Also, if $E$ is a topological-vectorial subspace of the Banach space $F$, and if the canonical map $E\hotimes G\to F\hotimes G$ is a topological isomorphism when $G=F'$, then $E''$ is a direct factor of $F''$;
  thus, in the frequent case where $E$ is already a direct factor of $E''$, $E$ will also be a direct factor of $F$.

  A useful case where the tensor product $u\hotimes v$ of two topological isomorphisms is a topological isomorphism is the following:
  \emph{If $E''$ is the bidual of $E$, then $E\hotimes F$ can be identified with a topological-vectorial subspace (resp. a normed vector subspace, if $E$ and $F$ are normed) of $E''\hotimes F$}.
}


\subsection{Nuclear maps}
\label{1.7}

(\cite[chap.~1, \S3, n\textsuperscript{o}~2]{PTT}).
\medskip

Let $E$ and $F$ be Banach spaces.
The continuous bilinear map $(x',y)\mapsto x'\otimes y$ from $E'\times F$ to $\LL(E,F)$ defines a natural continuous linear map from $E'\hotimes F$ to $\LL(E,F)$;
the elements of the image of $E'\hotimes F$ in $\LL(E,F)$ are said to be \emph{nuclear maps} from $E$ to $F$.
(We also define, between arbitrary locally convex spaces, the notions of \emph{trace maps} and \emph{Fredholm maps} --- see \hyperref[appendix1]{Appendix~1} --- which coincide with the notion of nuclear maps if $E$ and $F$ are Banach spaces; in this case, we can thus freely switch between saying ``nuclear maps'', ``trace maps'', or ``Fredholm maps''.)
Nuclear maps from $E$ to $F$ form a vector space which can be identified with a quotient space of $E'\hotimes F$ (and with $E'\hotimes F$ itself in all known cases --- see ``problem of bijectivity'' in \hyperref[1]{\S1}).
The quotient norm, again denoted by $u\mapsto\|u\|_1$, is called the \emph{trace norm} of the nuclear operator $u$.

If $E$ and $F$ are arbitrary locally convex spaces, then a linear map $u$ from $E$ to $F$ is said to be \emph{nuclear} if it is the composition of a sequence of three operators
\[
  E \xrightarrow{\alpha} E_1 \xrightarrow{\beta} F_1 \xrightarrow{\gamma} F
\]
where $E_1$ and $F_1$ are Banach spaces, $\beta$ is a nuclear map from $E_1$ to $F_1$, and $\alpha$ and $\gamma$ are continuous linear maps.
It is
\oldpage{86}
equivalent to say that there exists an weakly closed, equicontinuous, disked subset $A$ of $E'$, and a bounded disked subset $B$ of $F$ such that $F_B$ is complete, and finally some $u_0\in E'_A\hotimes F_B$ such that $u$ is the operator from $E$ to $F$ defined by $u_0$ (\emph{a priori}, $u_0$ defines a nuclear map from $\widehat{E}_{A^0}$ to $F_B$).
\emph{A nuclear map is always compact} (i.e. sends a suitable neighbourhood of $0$ to a relatively compact set).
Even better: by the corollary of \hyperref[1.theorem2]{Theorem~2}, we can suppose (in the above) that $B$ and $A$ are \emph{compact} subsets of $F$ and strong $E'$ (respectively).
\hyperref[1.theorem2]{Theorem~2}, applied directly, also gives the following: \emph{nuclear maps from $E$ to $F$ are exactly the maps that are sums of series (always absolutely convergent in $\LL(E,F)$ endowed with the bounded-convergence topology) $u=\sum\lambda_i x'_i\otimes y_i$, where $(x'_i)$ is an equicontinuous sequence in $E'$, $(y_i)$ is a \todo a compact disk of $F$, and $(\lambda_i)$ is a summable sequence of scalars}.
If we compose a nuclear map, on the left or on the right, with a continuous linear map, then we obtain another nuclear map.
The dual of a nuclear map from $E$ to $F$ is a nuclear map from strong $F'$ to strong $E'$ (and even from $F'$ endowed with the uniform convergence on compact disks of $F$, to strong $E'$).

Nuclear maps from $E$ to itself (more precisely, the slightly larger category of Fredholm maps from $E$ to itself) form a natural domain for \emph{Fredholm theory}.
Here, our interest lies in other properties of these operators, that result directly from either \hyperref[1.theorem2]{Theorem~2} or the corollary of \hyperref[1.theorem5]{Theorem~5}.

\begin{itenv}{Theorem 6}
\label{1.theorem6}
  Let $E$ and $G$ be locally convex spaces, and $F$ a vector subspace of $E$.
  Then:
  \begin{enumerate}[a)]
    \item Every nuclear map from $F$ to $G$ is the restriction of a nuclear map from $E$ to $G$.
    \item Suppose that $F$ is closed, and that every compact disc of $E/F$ is contained in the canonical image of a bounded disc $A$ of $E$ such that $E_A$ is complete (for example, a \emph{complete} bounded disc).
      (It suffices, for example, for $E$ to be an $\FF$-space, or for it be the dual of an $\FF$-space and for $F$ to be \emph{weakly} closed.)
      Then every nuclear map from $G$ to $E/F$ can be obtained from a nuclear map from $G$ to $E$ by passing to the quotient.
  \end{enumerate}
\end{itenv}

We have analogous statements for ``equinuclear'' sets of maps, if by that we mean a set of maps from a locally convex space $M$ to another locally convex space $N$, contained in
\oldpage{87}
the set of maps defined by the unit ball of a space $M'_A\hotimes N_B$, where $A$ is a weakly closed equicontinuous disk of $M'$, and $B$ a bounded disk in $N$ such that $N_B$ is complete.

We note that, if $u$ is a nuclear map from $E$ to $F$, $M$ a closed vector subspace of $E$ contained in the kernel of $u$, and $N$ a closed vector subspace of $F$ containing $u(E)$, then, in general, the linear map from $E/M$ to $F$ or from $E$ to $N$ defined by $u$ \emph{is not nuclear}, even if $E$ and $F$ are reflexive Banach spaces.
However, if $M$ (resp. $N$) admits a topological complement, then the map from $E/M$ to $F$ (resp. from $E$ to $N$) defined by $u$ will itself be nuclear.
This is the case, in particular, if $E$ (resp. $F$) is a Hilbert space.


\subsection{Integral linear maps, integral bilinear forms}
\label{1.8}

(\cite[chap.~1, \S4, n\textsuperscript{o}~3 and n\textsuperscript{o}~4]{PTT}).
\medskip

\begin{itenv}{Theorem 7}
\label{1.theorem7}
  Let $E$ and $F$ be two locally convex (resp. normed) spaces, and $v$ a separately continuous bilinear form on $E\times F$.
  Then the following conditions are equivalent:
  \begin{enumerate}[a)]
    \item $u\mapsto\langle u,v\rangle$ is a linear form on $E\otimes F$ that is continuous for the topology induced by $E\hhotimes F$ (resp. $u\mapsto\langle u,v\rangle$ is a linear form on $E\otimes F$ of norm~$\leq1$ when $E\otimes F$ is endowed with the norm induced by $E\hhotimes F$).
    \item $v$ is contained in the closed disked hull in $\sBB_\mathrm{s}(E,F)$ (the space $\sBB(E,F)$ endowed with the simple-convergence topology) of a set $M\otimes N$, where $M$ is an equicontinuous subset of $E'$, and $N$ an equicontinuous subset of $F'$ (resp. $M$ the unit ball of $E'$, and $N$ the unit ball of $F'$).
    \item There exists a measure $\mu$ on the product space of a weakly compact equicontinuous subspace $M$ of $E'$ with a weakly compact equicontinuous subspace $N$ of $F'$ (resp. a measure $\mu$ of norm~$\leq1$ on the product of the unit ball of $E'$ with the unit ball of $F'$, endowed with the product of the weak topologies) such that we have the formula
      \[
        v = \int_{M\times N} x'\otimes y' \dd\mu(x',y')
      \]
      (the weak integral in $\sBB(E,F)$, in duality with $E\otimes F$).
    \item There exists a compact space endowed with a positive measure $\mu$ of norm~$\leq1$, a continuous linear map $\alpha$ from $E$ to $\LL^\infty(\mu)$, and a continuous linear map $\beta$ from $F$ to $\LL^\infty(\mu)$ (resp. the same, but with $\alpha$ and $\beta$
\oldpage{88}
      also of norm~$\leq1$) such that we have $u(x,y)=\langle\alpha x,\beta y\rangle$ for $x\in E$, $y\in F$.
  \end{enumerate}
\end{itenv}

A bilinear form on $E\times F$ is said to be \emph{integral} if it satisfies any of the equivalent conditions of \hyperref[1.theorem7]{Theorem~7}.
In particular, the dual of $E\hhotimes F$ can be identified with the space $\JJ(E,F)$ of integral bilinear forms on $E\times F$.
If $E$ and $F$ are Banach spaces, then $\JJ(E,F)$ will be endowed with the dual norm of the Banach space $E\hhotimes F$, which we call the \emph{integral norm}, and denote by $\|v\|'_1$.
Similarly, a linear map $v$ from one locally convex space $E$ to another $G$ is said to be \emph{integral} if the corresponding bilinear form on $E\times G'$ is integral.

If $E$ and $G$ are Banach spaces, then we also call the integral norm of the corresponding bilinear form of $v$ the integral norm of $v$.

Recall that, in all known cases, when $E$ and $F$ are Banach, the natural linear map from $E'\hotimes F'$ to $\JJ(E,F)$ is a metric isomorphism from the first space to the second (see \hyperref[1.5]{\S5}), which is why we use the notation $\|v\|'_1$ for the integral norm, closely related to the notation $\|v\|_1$ for the trace norm.
Criterion~d) of \hyperref[1.theorem7]{Theorem~7} takes the following form (which we state for Banach spaces, for simplicity) for integral linear maps:
\emph{Let $v$ be a linear map from one Banach space $E$ to another $F$. Then $v$ is integral and of integral norm $\leq1$ if and only if the map from $E$ to $F''$ that it defines can be obtained by composing a linear map of norm $\leq$ from $E$ to some space $\LL^\infty(\mu)$, constructed with a suitable positive measure of norm $\leq1$ on a compact space, with the identity map from $\LL^\infty(\mu)$ to $\LL^1(\mu)$, and finally with a linear map of norm $\leq$ from $\LL^1(\mu)$ to $F''$.}
Similarly, criterion~b) of \hyperref[1.theorem7]{Theorem~7} easily gives:
\emph{The linear map $v$ from $E$ to $F$ is integral and of integral norm $\leq1$ if and only if it is an adherent point in $\LL_s(E,F_s)$ (where $F_s$ denotes $F$ endowed with the weak topology, and $\LL_s(E,F_s)$ denotes $\LL(E,F_s)$ endowed with the simple-convergence topology) of the disked hull of the set of the $x'\otimes y$, where $x'$ (resp. $y$) runs over the unit ball of $E'$ (resp. of $F$); or if it is adherent to the set of nuclear operators of trace-norm $\leq1$.}

\begin{rmenv}{Examples}
  Let $E$ and $F$ be arbitrary locally convex spaces.
  Then every bilinear form on $E\times F$ defined by an element of a space $E'_A\hotimes F'_B$, where $A$ (resp. $B$) is a weakly closed disked equicontinuous subset of $E'$ (resp. of $F'$), is integral.
  Thus every nuclear map from $E$ to $F$ is integral.
  The
\oldpage{89}
  converse is false, even for Banach spaces, since the identity map from $\LL^\infty(\mu)$ to $\LL^1(\mu)$ is integral, but it is not, in general, even compact.
  If $E=\CC(M)$ (resp. $F=\CC(N)$) is the space of continuous scalar functions on the compact space $M$ (resp. $N$), then we have seen (\hyperref[1.theorem4]{Theorem~4}) that $\CC(M)\hotimes\CC(N)$ can be identified with its norm to the space $\CC(M\times N)$, and so the space of integral bilinear forms on $\CC(M)\times\CC(N)$ can be identified with its norm to the space of Radon measures on the compact space $M\times N$.
  Other examples will be seen in \hyperref[1.9]{\S9}

  By composing an integral linear map on the left or on the right with a continuous linear map, we obtain another integral linear map.
  The transpose of an integral map from $E$ to $F$ is an integral map from strong $F'$ to strong $E'$.

  Using, for example, criterion~a) of \hyperref[1.theorem7]{Theorem~7}, we see that, if $E_1$ and $E_2$ are locally convex spaces, and $F_1$ (resp. $F_2$) is a topological vector subspace of $E_1$ (resp. of $E_2$), then every integral bilinear form on $F_1\times F_2$ can be extended to a integral bilinear form on $E_1\times E_2$, with equal integral norm if $E_1$ and $E_2$ are normed (compare with \hyperref[1.theorem6]{Theorem~6}).
  The most important properties of integral maps are summarised in the following:
\end{rmenv}

\begin{itenv}{Theorem 8}
\label{1.theorem8}
  Let $u$ be an integral linear map from a locally convex space $E$ to a locally convex space $F$.
  \begin{enumerate}
    \item If $F$ is quasi-complete, then $u$ is weakly compact, and sends weakly compact subsets of $E$ to compact subsets of $F$.
      If $v$ is a linear map from $F$ to a locally convex space $G$ that sends bounded subsets to weakly relatively compact subsets, then $v\circ u$ is a compact map.
    \item Let $v$ be a linear map from $F$ to an $\FF$-space $G$ that sends bounded subsets to weakly relatively compact subsets (resp. a linear map from a $\DF$-space $G$ to $E$ that sends bounded subsets to weakly relatively compact subsets of $E$).
      Then $v\circ u$ (resp. $u\circ v$) is a \emph{nuclear} map from $E$ to $G$ (resp. from $G$ to $F''$).
      If $E$, $F$, and $G$ are Banach spaces, then $\|v\circ u\|'_1\leq\|v\|\|u\|'$.
  \end{enumerate}
\end{itenv}

\begin{itenv}{Corollary 1}
\label{1.corollary1}
  The composition of two integral maps is a nuclear map.
\end{itenv}

\begin{rmenv}{Other corollaries}
  An integral map from $E$ to $F$ is nuclear if $F$ is a reflexive $\FF$-space, and it is
\oldpage{90}
  a nuclear map from $E$ to $F''$ if $E$ is a reflexive $\DF$-space.
  An integral map from $E$ to $F$ sends summable sequences to absolutely summable sequences, and weakly convergent sequences to nuclearly convergent sequences (see the end of \hyperref[1.4]{\S4}).
\end{rmenv}

\aster{
  So, considering the unit circumference $T$ of the complex plane, endowed with its Haar measure $\mu$, suppose that the sequence $(a_n)$ in the set $\mathbb{Z}$ of integers is such that $(\varepsilon_n a_n)$ is the sequence of coefficients of a function in $\LL^\infty(\mu)$, for some sequence $(\varepsilon)$ of numbers equal to $+1$ or $-1$;
  then we can easily see that the sequence $a_nz^n\in\LL^\infty(\mu)$ is summable, and is thus an absolutely summable sequence in $\LL^1(\mu)$, whence, immediately, $(a_n)\in\ll^1$.
  We have thus obtained an analogue of the well-known theorem of Littlewood (in which $\LL^\infty$ is replaced by $\LL^1$, and $\ell^1$ by $\ell^2$).
  We note that Littlewood's theorem can be obtained in the same way, as a consequence of the following theorem, which has much more general consequences (and which will later be published, along with some of its various consequences):
  \emph{every summable sequence in an $\LL^1(\mu)$-space (for an arbitrary measure) has a sequence of norms that is square summable} (and even belonging to $\ell^2\hotimes\LL^1(\mu)$;
  this is dual to the theorem, noted at the end of \hyperref[1.4]{\S4}, that says that every square summable sequence in a $\CC(K)$-space --- and even every sequence belonging to $\ell^2\hhotimes\CC(K)$ --- is nuclearly convergent to $0$).

  We will give some hints concerning the proof of \hyperref[1.theorem8]{Theorem~8}, that rely in an essential manner on criterion~d) of \hyperref[1.theorem7]{Theorem~7}.
  The fact that $u$ is a weakly compact map follows immediately from the fact that the identified map from $\LL^\infty(\mu)$ to $\LL^1(\mu)$ is a weakly compact map.
  The other claims of the theorem follow easily from the second part, which is more difficult.
  We can reduce to showing that every weakly compact linear map from $\LL^1(\mu)$ to an $\FF$-space $G$ induces a nuclear map from $\LL^\infty(\mu)$ to $G$, and thus (\hyperref[1.theorem3]{Theorem~3}) comes from some $f\in\LL^1_G(\mu)$.
  But a theorem of Dunford--Pettis--Phillips tells us that such a map is in fact given by a \emph{strongly measurable and bounded} map $f$ from $M$ to $G$.
  Note that \hyperref[1.corollary1]{Corollary~1}, which is important due to its application in the theory of nuclear spaces, admits a simpler direct proof:
  we can reduce to showing that, if $\mu$ and $\nu$ are measures on compact $M$ and $N$, then a continuous linear map $u$ from $\LL^1(\mu)$ to $\LL^\infty(\nu)$ defines a \emph{nuclear} map from $\LL^\infty(\mu)$ to $\LL^1(\nu)$.
  But $u$ can be identified with a continuous linear form on $\LL^1(\mu)\hotimes\LL^1(\nu)$, which is isomorphic to the space $\LL^1(\mu\otimes\nu)$ (\hyperref[1.theorem3]{Theorem~3}), and so $u$ is defined
\oldpage{91}
  by a measurable and bounded function $f$ on $M\times N$, which can be identified a fortiori with an element of $\LL^1(\mu\otimes\nu)=\LL^1(\mu)\hotimes\LL^1(\nu)$.
  This latter element indeed defines a nuclear map from $\LL^\infty(\mu)$ to $\LL^1(\nu)$, which happens to be exactly the map induced by $u$.
  QED.

  In \cite{PTT}, we deduce \hyperref[1.theorem8]{Theorem~8} from more general results (see \cite[chap.~1, \S4, n\textsuperscript{o}~2]{PTT}).
  The majority of \cite[chap.~1, \S4]{PTT} (the densest part of the whole work) is dedicated to the exposition of these results and their many consequences, which we cannot give in this summary.
}


\subsection{Integral linear maps to an \texorpdfstring{$\LL^1$}{L1} space or a \texorpdfstring{$\CC_0(M)$}{C0(M)} space}
\label{1.9}

(\cite[chap.~1, \S4, n\textsuperscript{o}~4]{PTT}).
\medskip

\aster{
  We can characterise the integral linear from a locally convex space $E$ to an $\LL^1(\mu)$ space (where $\mu$ is an arbitrary measure on a locally compact space $M$): they are the linear maps that send a suitable neighbourhood $V$ of $0$ in $E$ to a \emph{latticially bounded} subset of $\LL^1(\mu)$.
  If $E$ is normed, with $V$ its unit ball, and $h=\sup_{x\in V}|ux|$ (so that $h$ is a positive element of $\LL^1(\mu)$), then $\|u\|'_1=\|h\|_1$.
  If $E$ or $\LL^1(\mu)$ is separable, then the theorem of Dunford--Pettis gives an equivalent criterion (which we state, as an example, in the case where $E$ is assumed to be normed):
  \emph{there exists a weakly measurable map $f$ from $M$ to $E'$ such that $\|f(t)\|$ is a summable function of $t$, and such that, for all $x\in E$, $ux$ is the class in $\LL^1(\mu)$ of the function $t\mapsto\langle x,f(t)\rangle$; then $\|u\|'_1\leq\int\|f(t)\|\dd\mu(t)$ (and we have equality for a suitable choice of $f$).}
  We can also characterise the \emph{nuclear} maps from a locally convex space $E$ to $\LL^1(\mu)$: they are those which send a suitable neighbourhood $V$ of $0$ in $E$ to a subset $A$ of $\LL^1(\mu)$ which is latticially bounded and further \emph{equimeasurable} (by which we mean that, for every compact $K\subset M$, and every $\varepsilon>0$, there exists a compact $K_0$ such that $\mu(K\cap K_0^c)\leq\varepsilon$, and such that the $\varphi\in A$ agree almost everywhere on $K_0$ with the functions from an equicontinuous and uniformly bounded set of functions on $K_0$).
  If we suppose, for simplicity, that $E$ is a Banach space, then this is equivalent to saying that the map in question is given by an \emph{integrable} \cite{2} map $f$ from $M$ to $E'$ (and, indeed, it follows immediately from \hyperref[1.theorem3]{Theorem~3} that this indeed implies that $u$ is nuclear).

  Dually, let $M$ be a locally compact space, and $E$ a Banach space (for simplicity), and suppose that either $M$ is metrisable and
\oldpage{92}
  countable at infinity, or that $E$ is separable.
  Then the integral maps $u$ from $\CC_0(M)$ (the space of continuous functions on $M$ that are ``zero at infinity'') to $E'$, i.e. the continuous linear forms on $\CC_0(M,E)=\CC_0(M)\hhotimes E$, are exactly those given by a measure $\mu$ on $M$ and a weakly $\mu$-measurable and bounded map $f$ from $M$ to $E'$ by the formula $u\varphi=\int\varphi(t)f(t)\dd\mu(t)$.
  We can further suppose that $\|f(t)\|=1$ for all $t$, and that $\|\mu\|_1=\|u\|'_1$.
  Using \hyperref[1.theorem3]{Theorem~3}, we also find that the \emph{nuclear} maps from $\CC_0(M)$ to $E'$, or even from $\CC_0(M)$ to an arbitrary Banach space $F$, are exactly those given by a pair $(u,f)$ as above, but with $f$ being an \emph{integrable} map from $M$ to $F$;
  it is no longer necessary here to make any separability hypotheses.
  In particular, if $K$ and $L$ are compact spaces, $\mu$ a measure on $K$, and $N(x,y)$ a continuous scalar function on $K\times L$, then the map $f\mapsto\int f(x)N(x,y)\dd\mu(x)$ from $C(K)$ to $C(L)$, defined by the continuous kernel $N$, is nuclear.
  Note that we have just described two remarkable categories of ``vectorial measures'' on $M$, which we call \emph{integral vectorial measures} and \emph{nuclear vectorial measures} (respectively) on $M$.

  The fact that we can characterise the integral and nuclear maps from a Banach space (for example) $E$ to $\LL^1(\mu)$ by properties concerning their images of the unit ball is completely special to $\LL^1(\mu)$ (see the end of \hyperref[1.7]{\S7}) and is also linked to the corollary of \hyperref[1.theorem3]{Theorem~3}.
  Similarly, we can show that the integral linear maps from $\LL^1(\mu)$ to a locally convex space $E$ can be characterised by properties concerning their images of the unit ball of $\LL^1(\mu)$, a feature that we will not further study here (see \cite[chap.~1, \S4, n\textsuperscript{o}~6]{PTT}).
}


\subsection*{Appendix 1}
\label{appendix1}
\addcontentsline{toc}{subsection}{Appendix 1}




%% Bibliography %%

\nocite{*}

\begin{thebibliography}{9}

  \bibitem{1}
  {Bourbaki, N.}
  \newblock {\em Alg\`{e}bre multilin\'{e}aire}.
  \newblock Hermann, {\em Act. Sc. Ind.} \textbf{1044}.

  \bibitem{2}
  {Bourbaki, N.}
  \newblock {\em Int\'{e}gration}.
  \newblock Hermann, {\em Act. Sc. Ind.} \textbf{1175}.

  \bibitem{3}
  {Dieudonn\'{e}, J. and Schwartz, L.}
  \newblock La dualit\'{e} dans les espaces $\FF$ et $\DF$.
  \newblock {\em Annales de Grenoble} \textbf{1} (1949), 61--101.

  \bibitem{4}
  {Dixmier, J.}
  \newblock Les fonctionnelles lin\'{e}aires sur l'ensemble des op\'{e}rateurs born\'{e}s d'un espace de Hilbert.
  \newblock {\em Annals of Math.} \textbf{51} (1950), 387--408.

  \bibitem[PTT]{PTT}
  {Grothendieck, A.}
  \newblock Produits tensoriels topologiques et espaces nucl\'{e}aires
  \newblock {\em Memoirs of the Amer. Math. Soc.}, to appear.

  \bibitem{6}
  {Grothendieck, A.}
  \newblock Sur les espaces $\FF$ et $\DF$.
  \newblock {\em Summa Brasiliensis Mathematicae}, to appear.

  \bibitem{7}
  {K\"{o}the, G.}
  \newblock Die Stufenr\"{a}ume, eine einfache klasse linearer vollkommener R\"{a}ume.
  \newblock {\em Math. Zeitschrift} \textbf{51} (1948), 317--345.

  \bibitem{8}
  {Schatten, R.}
  \newblock {\em A theory of cross-spaces}.
  \newblock Princeton University Press (1950).

  \bibitem{9}
  {Schwartz, L.}
  \newblock {\em Th\'{e}orie des distributions}, t.~1 and 2.
  \newblock Hermann, {\em Act Sc. et Ind.} \textbf{1091} and \textbf{1122}.

\end{thebibliography}


\end{document}
