\documentclass{article}

\title{Singularities of differentiable maps}
\author{A. Haefliger}
\date{13\textsuperscript{th} and 24\textsuperscript{th} of May, 1957}

\usepackage{amssymb,amsmath}

\usepackage{hyperref}
\usepackage{xcolor}
\hypersetup{colorlinks,linkcolor={red!50!black},citecolor={blue!50!black},urlcolor={blue!80!black}}
\usepackage[nameinlink]{cleveref}
\usepackage{enumerate}
\usepackage{tikz-cd}

\usepackage{mathrsfs}
%% Fancy fonts --- feel free to remove! %%
\usepackage{fouriernc}


\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{xstring}
\makeatletter
\ifx\pdfmdfivesum\undefined
  \let\pdfmdfivesum\mdfivesum
\fi
\edef\filesum{\pdfmdfivesum file {\jobname}}
\pagestyle{fancy}
\makeatletter
\let\runauthor\@author
\let\runtitle\@title
\makeatother
\fancyhf{}
\lhead{\footnotesize\runtitle}
\lfoot{\footnotesize Version: \texttt{\StrMid{\filesum}{1}{8}}}
\cfoot{\small\thepage\ of \pageref*{LastPage}}


\crefname{section}{\S\!}{\S\S\!}
\crefname{equation}{}{}


%% Theorem environments %%

\usepackage{amsthm}

\theoremstyle{plain}
  
  \newtheorem*{lemma*}{Lemma}


\theoremstyle{definition}

  \newtheorem*{remarks*}{Remarks}


%% Shortcuts %%

\newcommand{\RR}{\mathbb{R}}

\renewcommand{\geq}{\geqslant}
\renewcommand{\leq}{\leqslant}

\newcommand{\oldpage}[1]{\marginpar{\footnotesize$\Big\vert$ \textit{p.~#1}}}


%% Document %%

% \usepackage{embedall}
\begin{document}

\maketitle
\thispagestyle{fancy}

\renewcommand{\abstractname}{Translator's note.}

\begin{abstract}
  \renewcommand*{\thefootnote}{\fnsymbol{footnote}}
  \emph{This text is one of a series\footnote{\url{https://thosgood.com/translations/}} of translations of various papers into English.}
  \emph{The translator takes full responsibility for any errors introduced in the passage from one language to another, and claims no rights to any of the mathematical content herein.}

  \medskip
  
  \emph{What follows is a translation of the French seminar talk:}

  \medskip\noindent
  \textsc{Haefliger, A.}
  ``Les singularit\'{e}s des applications diff\'{e}rentiables''.
  \emph{S\'{e}minaire Henri Cartan}, Volume~\textbf{9} (1956--1957), Talk no.~7.
  {\url{http://www.numdam.org/item/SHC_1956-1957__9__A7_0/}}
\end{abstract}

\setcounter{footnote}{0}

\tableofcontents
\bigskip


%% Content %%

\oldpage{7-01}
The aim of this expos\'{e} is to reformulate, in the terminology of ``jets'' of C.~Ehresmann, the basis of the theory of singularities of differentiable maps developed by H.~Whitney and R.~Thom.
This language seems to us to be well adapted to make precise the definitions and the problems, as well as to bring out their invariant (under a change of coordinates) characteristics.
This first talk contains almost only the definitions needed to be able to state the Thom lemma and the fundamental conjecture (Whitney--Thom) of the theory.
The following talk will discuss the question of homological invariance of critical cycles.

We refer to \cite{2} and \cite{7} for examples.


\section{Recalling some definitions}
\label{section1}

We briefly recall here some notation and definitions of the theory of jets of C.~Ehresmann.
For more details, see \cite{1}.

We write ``$r$-differentiable map'' to mean a differentiable map of class~$r$.
We denote by $\RR^n$ the real space of dimension~$n$.

In the set of pointed local $r$-differentiable maps from $\RR^n$ to $\RR^p$ (that is, the set of pairs $(f,x)$ consisting of an $r$-differentiable map $f$ from an open subset $U$ of $\RR^n$ to $\RR^p$, and of a point $x\in U$), we say that the pair $(f,x)$ is equivalent to the pair $(f',x')$ if and only if $x=x'$ and all the partial derivatives of order $\leq r$ of $f$ and $f'$ at the point $x$ are equal.
The equivalence class of the pair $(f,x)$ is called the \emph{jet of order~$r$ of $f$ at the point $x$}, and is denoted by $f^r(x)$ (instead of $j_x^rf$).
The point $x$ is said to be the \emph{source} of the jet $f^r(x)$, and $f(x)$ is its \emph{target}.
If $V^n$ and $V^p$ are $r$-differentiable manifolds, then we similarly have the notion of a jet of order~$r$ at a point $x$ of $V^n$ of a local $r$-differentiable map from $V^n$ to $V^p$.

The set $J^r(\RR^n,\RR^p$ of jets of order $r$ (of $r$-differentiable maps) from $\RR^n$ to $\RR^p$ is endowed with a topological structure which makes it a manifold homeomorphic to some real space:
the coordinates of a jet $f^r(x)$ are the values of the coordinates of its source $x$ and of its target $f(x)$ and of the partial derivatives up to order~$r$ of $f$ at the point $x$.
The space of jets of order~$r$ from
\oldpage{7-02}
$\RR^n$ to $\RR^p$ with source and target being the origin of $\RR^n$ and $\RR^p$ (respectively) is denoted by $L_{n,p}^r$.
The space $J^r(\RR^n,\RR^p)$ is canonically isomorphic to the product $\RR^n\times\RR^p\times L_{n,p}^r$.

The partial composition law defined between pointed local maps from $\RR^n$ to $\RR^p$ and pointed local maps from $\RR^p$ to $\RR^m$ defines, by passing to the quotient, a partially-defined \emph{composition law} between $J^r(\RR^n,\RR^p)$ and $J^r(\RR^p,\RR^m)$ that is continuous.
The composition $YX$ of $X\in J^r(\RR^n,\RR^p)$ and $Y\in J^r(\RR^p,\RR^m)$ is defined if the source of $Y$ is the target of $X$.

The set of \emph{invertible jets of order~$r$ from $\RR^n$ to $\RR^n$} with source and target equal to the origin forms an algebraic \emph{group} denoted by $L_n^r$.
\emph{The product group $L_n^r\times L_p^r$ acts as a transformation group on $L_{n,p}^r$}: the transformation of an element $X\in L_{n,p}^r$ under the pair $(Z,Z')\in L_n^r\times L_p^r$ is the element $Z'XZ^{-1}$.

All of the above notions immediately generalise if we replace $\RR^n$ and $\RR^p$ by $r$-differentiable manifolds $V^n$ and $V^p$ (of dimension~$n$ and $p$, respectively).
The space $J^r(V^n,V^p)$ of jets of order~$r$ from $V^n$ to $V^p$ is endowed with the structure of a manifold;
the map that sends any element of $J^r(V^n,V^p)$ to the pair consisting of its source and target is the projection of a fibration with base $V^n\times V^p$, with fibre $L_{n,p}^r$, and with structure group $L_n^r\times L_p^r$.

Every $r$-differentiable map $f$ from $V^n$ to $V^p$ \emph{extends} to a map $f^r$ from $V^n$ to $J^r(V^n,V^p)$ that sends any point $x$ of $V^n$ to $f^r(x)$ (i.e. the jet of order~$r$ of $f$ at the point $x$).

A jet $X\in J^r(V^n,V^p)$ is said to be \emph{equivalent} to a jet $X'\in J^r(V'^n,V'^p)$ (written $X\sim X'$) if there exist invertible jets $Z_1\in J^r(V^n,V'^n)$ and $Z_2\in J^r(V^p,V'^p)$ such that $X'=Z_2XZ_1^{-1}$.
In particular, two jets $X$ and $X'$ belonging to $L_{n,p}^r$ are equivalent if and only if there exists an elements of the group $L_n^r\times L_p^r$ that sends $X$ to $X'$.

Finally, note that the space $J^{r+k}(V^n,V^p)$ can be identified with a sub-manifold of the space of jets of order~$k$ from $V^n$ to $J^r(V^n,V^p)$.
This is a direct consequence of the fact that the partial derivatives of order $\leq r+k$ of an $(r+k)$-differentiable map from $\RR^n$ to $\RR^p$ are the partial derivatives of order $\leq k$ of the partial derivatives of order $\leq r$ of $f$.


\section{The Thom lemma}
\label{section2}

\subsection*{The notion of a map transversal to a sub-manifold}

\oldpage{7-03}

An \emph{$r$-differentiable sub-manifold $S$ of codimension~$q$} of an $r$-differentiable manifold $V^p$ is a subset $S$ of $V^p$ such that, for every point $y$ of $S$, there exists a neighbourhood $U_y$ of $y$, and an $r$-differentiable map $g_y$ from $U_y$ to $\RR^q$ of rank~$q$ such that $U_y\cap S$ is the inverse image under $g_y$ of the origin $0$ in $\RR^q$.

A differentiable map $f$ from a differentiable manifold $V^n$ to $V^p$ is said to be \emph{transversal to the sub-manifold $S$ at the point $x$ of $V^n$} if either
\begin{enumerate}
  \item $f(x)\not\in S$, or
  \item $f(x)\in S$, and the image under $f$ of the tangent space of $V^n$ at the point $x$ and the tangent space of $S$ at the point $f(x)$ generate the tangent space of $V^p$ at the point $f(x)=y$.
    (This is equivalent to saying that $g_y\circ f$ is a differentiable map of rank~$q$ at the point $x$, which is a property concerning the first-order jet of $f$ at the point $x$.)
\end{enumerate}

If $f$ (assumed to be $r$-differentiable) is transversal to $S$ at every point of $V^n$, then \emph{$f^{-1}(S)$ is an $r$-differentiable sub-manifold of $V^n$ of codimension~$q$}.
Indeed, $f^{-1}(S)$ is defined by $g_y\circ f=0$ in $f^{-1}(U_y)$.

A \emph{Collection of $r$-differentiable sub-manifolds} in $V^p$ is a union $S=S_1\cup S_2\cup\ldots\cup S_m$ of a finite number (with $m\leq p$) of disjoint $r$-differentiable sub-manifolds of $V^p$ such that
\begin{enumerate}
  \item $\dim S_k>\dim S_{k+1}$ ; and
  \item $S_k\cup S_{k+1}\cup\ldots\cup S_m$ is closed in $V^p$ (for $k=1,2\ldots,m$).
\end{enumerate}
The points of $S_1$ are said to be the \emph{regular} points of $S$.

An $r$-differentiable map $f$ from $V^n$ to $V^p$ is said to be transversal at a point $x$ of $V^n$ to the collection $S$ of manifolds if $f$ is transversal at the point $x$ to the sub-manifold $S_k$ containing $f(x)$ of the decomposition $S$.
If $f$ is transversal to $S$ at every point of $V^n$, then $f^{-1}(S)$ is a collection of $r$-differentiable manifolds in $V^n$, i.e. $f^{-1}(S)=f^{-1}(S_1)\cup\ldots\cup f^{-1}(S_m)$.

Every algebraic variety can be expressed (in general, in multiple ways) as a collection of partial varieties (cf. \cite{8}).

We will write $L(\RR^n,\RR^p,q)$ to denote the functional space of $q$-differentiable maps from $\RR^n$ to $\RR^p$ endowed with the topology defined by the chart of the partial derivatives of order~$\leq q$ on the compact subsets of $\RR^n$.

\begin{lemma*}[The Thom lemma]
  Let $S$ be an $s$-differentiable closed sub-manifold of $J^r(\RR^n,\RR^p)$ (with $s\geq n$).
  Then the $q$-differentiable maps (for $q\geq r+n$) from $\RR^n$
\oldpage{7-04}
  to $\RR^p$ such that the extension $f^r$ of order~$r$ of $f$ is transversal to $S$ at every point of a compact subset of $\RR^n$ form an everywhere-dense open subset of $L(\RR^n,\RR^p,q)$.
\end{lemma*}

\begin{proof}
  See \cite{3}.
\end{proof}

\begin{remarks*}
  Since $L^(\RR^n,\RR^p,q)$ is a Baire space, the $q$-differentiable maps that are not transversal to $S$ at at least one point of $\RR^n$ form a subset whose complement is everywhere dense in $L(\RR^n,\RR^p,q)$;
  in other words, every $r$-differentiable map from $\RR^n$ to $\RR^p$ can be approximated as close as desired, as can its partial derivatives of order~$\leq q$, by an $r$-differentiable map $g$ that is transversal to $S$ at every point of $\RR^n$.
  This is still true even if $S$ is a non-closed sub-manifold, or if $S$ is a collection of sub-manifolds of $J^r(\RR^n,\RR^p)$.
\end{remarks*}

We can prove the following by passing from local to global:

\begin{lemma*}[The global Thom lemma]
  Let $V^n$ and $V^p$ be (paracompact) differentiable manifolds, and let $S$ be an $s$-differentiable sub-manifold of $J^r(V^n,V^p)$ (for $s\geq n$).
  The $q$-differentiable maps (for $q\geq r+n$) from $V^n$ to $V^p$ such that the extension $f^r$ of $f$ is transversal to $S$ at every point of a compact subset of $V^n$ form an everywhere-dense open subset of $L(V^n,V^p,q)$.
\end{lemma*}


\section{Critical variety}
\label{3}




%% Bibliography %%

\nocite{*}

\begin{thebibliography}{1}

  \bibitem{1}
  {\sc Ehresmann, C.}
  \newblock ``Introduction \`{a} la th\'{e}orie des structures infinit\'{e}simales et des pseudo-groups de Lie'', in
  \newblock {\em Colloques Internationaux du Centre National de la Recherche Scientifique \textbf{52}: G\'{e}om\'{e}trie diff\'{e}rentielle, Strasbourg 1953}.
  \newblock Paris, Centre National de la Recherche Scientifique (1953), 97--110.

  \bibitem{2}
  {\sc Thom, R.}
  \newblock Les singularit\'{e}s des applications diff\'{e}rentiables.
  \newblock {\em Ann. Inst. Fourier} \textbf{6} (1955--56), 43--87.

  \bibitem{3}
  {\sc Thom, R.}
  \newblock Un lemma sur les applications diff\'{e}rentiables.
  \newblock {\em Bol. Soc. Mat. Mexico} \textbf{2} (1956), 59--71.

  \bibitem{4}
  {\sc Whitney, H.}
  \newblock Differentiable manifolds.
  \newblock {\em Annals of Math.} \textbf{37} (1936), 645--680.

  \bibitem{5}
  {\sc Whitney, H.}
  \newblock The general type of singularity of a set of $2n-1$ smooth functions of $n$ variables.
  \newblock {\em Duke Math. J.} \textbf{10} (1943), 161--172.

  \bibitem{6}
  {\sc Whitney, H.}
  \newblock On singularities of mappings of Euclidean spaces. I. Mappings of the plane into the plane.
  \newblock {\em Annals of Math.} \textbf{62} (1955), 374--410.

  \bibitem{7}
  {\sc Whitney, H.}
  \newblock Singularities of mappings of Euclidean spaces.
  \newblock (To appear).

  \bibitem{8}
  {\sc Whitney, H.}
  \newblock Elementary structure of real algebraic varieties.
  \newblock (To appear in {\em Annals of Math.}).

\end{thebibliography}


\end{document}
